---
title: "Unit3HW_mmastrangeli"
author: "Mark Mastrangeli"
date: "9/17/2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/mmastran/SMU/MSDS/github/MSDS/unit03")
```
Live Session Unit 03 Assignment

##1.	 GitHub Cloning (20 points): Using Git, clone the following GitHub repository to your local machine: https://github.com/caesar0301/awesome-public-datasets.  In RMarkdown, please show the code (commented out, as it’s not R syntax) that you used to create a new directory, navigate to the appropriate directory, and clone the repository to it.  One Git command per line, please.

```{r, echo=TRUE}
#Making sure that my working directory is set correctly to my github syncronized directory
setwd("/Users/mmastran/SMU/MSDS/github/MSDS/unit03")
#Creating variable to store URL of awesome-public-datasets
UrlAddress <- "https://github.com/caesar0301/awesome-public-datasets"
#Git Command to clone awesome-public-datasets to my MSDS/unit03 Github repository
git2r::clone(url = UrlAddress, local_path = "/Users/mmastran/SMU/MSDS/github/MSDS/unit03/awesome-public-datasets")
```

##2.	Data Summary (20 points): 
#From this aforementioned cloned repo, please extract titanic.csv.zip.  To be clear, this does not have to be done in Git or command line.  
*a.	In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df.  This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status.
*b.	Output the respective count of females and males aboard the Titanic. Plot the frequency of females and males.  Be sure to give an accurate title and label the axes.
*c.	Please use one apply function (to review: swirl() modules 11, 12) to output the means of Age, Fare, and Survival.  Make sure the output is a real number for all three means.

```{r, echo=TRUE}
#Creating variable to store path of the cloned awesome-public-datasets repo. I am sure there is a fancier way to code navigation from the cloned directory root to the sub directory that contains the titanic.csv.zip file. Would be great to improve this in that way. 
awesome_public_data_archive <- "/Users/mmastran/SMU/MSDS/github/MSDS/unit03/awesome-public-datasets/Datasets/titanic.csv.zip"
#extract titanic.csv from archive
unzip(awesome_public_data_archive)
```
##2a. In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df.  This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status.
```{r, echo=TRUE}
##Read in titanic.csv using readcsv() and store in a data frame name 'titanic_data' then pass to the str() function for analysis of structure

#Store titanic.csv path in variable
titanic_csv_path <- "/Users/mmastran/SMU/MSDS/github/MSDS/unit03/titanic.csv"

#read in data from titanic.csv
titanic_data <- read.csv(titanic_csv_path)

#Confirm class of titanic_data and analyze Structure using str() function to identify column names 'variables', variance and number of observations.
str(titanic_data)

```
##3b.	Output the respective count of females and males aboard the Titanic. Plot the frequency of females and males.  Be sure to give an accurate title and label the axes.
```{r, echo=TRUE}

#Load the plyr library for analysis and counting
library(plyr)

#extract Sex column of data and put in a vector
titanic_data_sex <- titanic_data$Sex

#display the titanic passenger sex dataframe
titanic_data_sex

#Subset the vector into two vectors, male and female
#titanic_data_sex_male <- subset(titanic_data_sex, subset("male") == TRUE, select, drop = FALSE)
#titanic_data_sex_male

#Try using count function
count(titanic_data_sex)
#count returns unique factors and frequency but does not plot well

plot(titanic_data_sex, main = "Titanic Male & Female Frequency", xlab = "Sex", ylab = "Frequency")

```
## 3c.	Please use one apply function (to review: swirl() modules 11, 12) to output the means of Age, Fare, and Survival.  Make sure the output is a real number for all three means.
```{r, echo=TRUE}
#Using sapply() return the mean of the 'Age', 'Fare', 'Survived' variables. In order to nest a function in the sapply function, using list to return
#Had to add the na.rm argument in order to ignore NA entries in the data. 

lapply(list(titanic_data$Age, titanic_data$Fare, titanic_data$Survived), mean, na.rm=TRUE)

#Trying the sapply function
sapply(list(titanic_data$Age, titanic_data$Fare, titanic_data$Survived), mean, simplify = TRUE, USE.NAMES = TRUE, na.rm=TRUE)

```


4.	Function Building (30 points): You research sleep and just got your first data set.  Later, you’ll have another dataset with the same column names, so you want to create a helper function that you can analyze this dataset and the next.  Load sleep_data_01.csv (found at http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/index.html).  Questions 3A through 3D should be answered in function(x){}.  3E can be outside of the function. 
*a.	Create objects for the median Age, the minimum and maximum Duration of sleep, and the mean and standard deviation of the Rosenberg Self Esteem scale (RSES).  You may need to specify a few options like in Problem 2 and live session.
*b.	Create a data.frame object called report: it should consist of the median age, the RSES mean and standard deviation respectively divided by five (since there are five questions and these scores are summed), and the range of Duration (the statistical definition of range; it should be a single number.)
*c.	Change the column names of this data.frame to MedianAge, SelfEsteem, SE_SD, and DurationRange.
*d.	Round the report to at most 2 digits: leave this as the closing line to the function.
*e.	Finally, run the function on your sleep data to show the output.

```{r, echo=TRUE}
#Download the file sleep_data_01.csv from http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/index.html and store in a dataframe 'sleep_data

# Download and store data from sleep_data_01.csv and store in a dataframe 'sleep_data' 
sleep_data <- read.csv("http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv")
str(sleep_data)
```

##4a.	Create objects for the median Age, the minimum and maximum Duration of sleep, and the mean and standard deviation of the Rosenberg Self Esteem scale (RSES).  You may need to specify a few options like in Problem 2 and live session.
```{r, echo=TRUE}

sleep_data_age <- sleep_data$Age
sleed_data_duration <- sleep_data$Duration
sleep_data_RSES <- sleep_data$RSES

#sleep_data_age
#sleed_data_duration
#sleep_data_RSES

sleep_data_age_median <- median(sleep_data_age, na.rm=TRUE)
sleed_data_duration_min <- min(sleed_data_duration, na.rm=TRUE)
sleed_data_duration_max <- max(sleed_data_duration, na.rm=TRUE)
sleep_data_RSES_mean <- mean(sleep_data_RSES, na.rm=TRUE)
sleep_data_RSES_sd <- sd(sleep_data_RSES, na.rm=TRUE)

sleep_data_age_median
sleed_data_duration_min
sleed_data_duration_max
sleep_data_RSES_mean
sleep_data_RSES_sd

```
## 4b.	Create a data.frame object called report: 
#it should consist of the median age, the RSES mean and standard deviation respectively divided by five (since there are five questions and these scores are summed), 
#and the range of Duration (the statistical definition of range; it should be a single number.)
```{r, echo=TRUE}
#DivBy2 <- function(x){x/2}
report <- round(c(c(sleep_data_age_median, sleep_data_RSES_mean, sleep_data_RSES_sd)/5, range(sleep_data$Duration, na.rm = TRUE)))

report
```

##4c.	Change the column names of this data.frame to MedianAge, SelfEsteem, SE_SD, and DurationRange.
```{r, echo=TRUE}


```

##4d.	Round the report to at most 2 digits: leave this as the closing line to the function.
```{r, echo=TRUE}


```
##4e.	Finally, run the function on your sleep data to show the output.
```{r, echo=TRUE}

```



5.	FiveThirtyEight Data (30 points): Navigate on GitHub to https://github.com/rudeboybert/fivethirtyeight and read README.md. It will include everything you need. 
*a.	Install the fivethirtyeight package.
*b.	In the listing of Data sets in package ‘fivethirtyeight,’ assign the 22nd data set to an object ‘df.’
*c.	Use a more detailed list of the data sets to write out the URL in a comment to the related news story.
*d.	Using R command(s), give the dimensions and column names of this data frame.
##
## 5a.	Install the fivethirtyeight package.
```{r, echo=TRUE}



```

