---
title: "Unit3HW_mmastrangeli"
author: "Mark Mastrangeli"
date: "9/17/2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/mmastran/SMU/MSDS/github/MSDS/unit03")
```
Live Session Unit 03 Assignment

##Submission 
#ALL MATERIAL MUST BE KNITTED INTO A SINGLE, LEGIBLE, AND DOCUMENTED HTML DOCUMENT. Formatting can be basic, but it should be easily human-readable.  Unless otherwise stated, please enable {r, echo=TRUE} so your code is visible.

##1.	 GitHub Cloning (20 points): Using Git, clone the following GitHub repository to your local machine: https://github.com/caesar0301/awesome-public-datasets.  In RMarkdown, please show the code (commented out, as it’s not R syntax) that you used to create a new directory, navigate to the appropriate directory, and clone the repository to it.  One Git command per line, please.

```{r, echo=TRUE}
#Creating variable to store URL of awesome-public-datasets
UrlAddress <- "https://github.com/caesar0301/awesome-public-datasets"
#Git Command to clone awesome-public-datasets to my MSDS Github repository
git2r::clone(url = UrlAddress, local_path = "/Users/mmastran/SMU/MSDS/github/MSDS/unit03/awesome-public-datasets")
```

##2.	Data Summary (20 points): 
#From this aforementioned cloned repo, please extract titanic.csv.zip.  To be clear, this does not have to be done in Git or command line.  
*a.	In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df.  This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status.
*b.	Output the respective count of females and males aboard the Titanic. Plot the frequency of females and males.  Be sure to give an accurate title and label the axes.
*c.	Please use one apply function (to review: swirl() modules 11, 12) to output the means of Age, Fare, and Survival.  Make sure the output is a real number for all three means.

```{r, echo=TRUE}
#Creating variable to store URL of the cloned awesome-public-datasets repo
UrlAddress <- "https://github.com/caesar0301/awesome-public-datasets"
#Git Command to clone awesome-public-datasets to my MSDS Github repository
git2r::clone(url = UrlAddress, local_path = "/Users/mmastran/SMU/MSDS/github/MSDS/unit03/awesome-public-datasets")
```
