---
title: "CaseStudy01RMD"
author: "Mark Mastrangeli"
date: "10/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Parsing data from .csv into dataframes 'Beers' and 'Breweries'
```{r , echo=TRUE}
library(base)
library(tidyverse)
library(knitr)

Beers <- read_csv("Beers.csv", col_names = TRUE, trim_ws = TRUE, na = "NA")
#View(Beers)
Breweries <- read_csv("Breweries.csv", col_names = TRUE, trim_ws = TRUE, na = "NA")
#View(Breweries)
```

##Questions
*1.	How many breweries are present in each state?
```{r, echo=FALSE}
#table(Breweries$State)
View(table(Breweries$State))


```
2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r, echo=FALSE}
Breweries_Beers <- merge(Beers, Breweries, by.x = ("Brewery_id"), by.y = ("Brew_ID"), sort = TRUE)
#View(Breweries_Beers)

head(Breweries_Beers, 6)
tail(Breweries_Beers, 6)

```

3.	Report the number of NA's in each column.
```{r , echo=TRUE}
colSums(is.na(Breweries_Beers))
#Can we display this somehow?

```

4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r , echo=TRUE}


State_ABV_Median <- tapply(Breweries_Beers$ABV, Breweries_Beers$State, median, na.rm = TRUE)
State_ABV_Median


State_IBU_Median <- tapply(Breweries_Beers$IBU, Breweries_Beers$State, median, na.rm = TRUE)
State_IBU_Median
ggplot(State_IBU_Median, aes(x=))


#Make States a factor in order to color bars of chart by state
#State_IBU_Median$ <- factor(State_IBU_Median, levels = State_IBU_Median)
#Plot the Number of VAMC per state and order
#ggplot(VAMC_State_Count, aes(reorder(VAMC_State_Count$State, VAMC_State_Count$`Number of VAMC`), VAMC_State_Count$`Number of VAMC`)) + geom_bar(aes(fill=VAMC_State_Count$State), stat="identity") + ggtitle("Distribution of VA Medical Centers in continental US") + theme(axis.text.x=element_text(angle = 90, vjust = 1,hjust = ), legend.position="none") + scale_color_manual(values = LST) + labs(title="Number of Veterans Affairs Medical Centers by State", fill = "State", x= "State", y= "VAMC Count")
```
5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r , echo=TRUE}

MaxABV <- max(Breweries_Beers$ABV, na.rm = TRUE)
MaxABV
MaxIBU <- max(Breweries_Beers$IBU, na.rm = TRUE)
MaxIBU

#Display the Row from the Breweries.Beers data frame for which Breweries.Beers$ABV = MaxABV

#Display the Row from the Breweries.Beers data frame for which Breweries.Beers$IBU = MaxIBU

#Display the State with MaxABV

#Display the State with MaxIBU

```


6.	Summary statistics for the ABV variable.
```{r , echo=TRUE}
#Need to display a summary of our findings
```


7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r , echo=TRUE}
ABV_State <- tapply(Breweries_Beers$ABV, Breweries_Beers$State, median, na.rm = TRUE)
ABV_State
IBU_State <- tapply(Breweries_Beers$IBU, Breweries_Beers$State, median, na.rm = TRUE)
IBU_State
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
