---
title: "CaseStudy2_final_mmastrangeli"
author: "Mark Mastrangeli"
date: "12/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(base)
library(class)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(knitr)
library(mlr)
library(caret)
#colnames(read.csv("CaseStudy2-data.csv", header = TRUE))
```
#Run KNN multiple times on individual variables to get  accuracy estimate. 
#Bivin Sadler: 2. Choose the top three variables with the highest accruacy.
#Bivin Sadler: 3. Now allow more than one variable in the knn model to get a more optimal accuracy on the dfVal  set. 
#Bivin Sadler: 4. Make some plots of the variables you chose versus attrition RATE.
#Bivin Sadler: Now for leraning about Job Role relationships ...
#Bivin Sadler: Plot Job role (x axis) versus satisfaction variables.  Enviromnment, etc. 
#Bivin Sadler: Then make a video and clean up the RMD file.  


```{r Import}
#Collecting and setting up persistent variables
dfTrain <- read_csv("CaseStudy2-data.csv")
dfVal <- read_csv("CaseStudy2Validation.csv")



dfTrainImportSum <- summary(dfTrain)
dfValImportSum <- summary(dfVal)
str(dfTrain)
str(dfVal)
head(dfTrain)

#Drop unused variables
dfTrain <- subset(dfTrain, select = -c(EmployeeCount, EmployeeNumber, Over18, Rand, StandardHours))
dfValTrue <- subset(dfVal, select = Attrition)
dfVal <- subset(dfVal, select = -c(EmployeeCount, EmployeeNumber, Over18, Rand, StandardHours))
dfValTrue <- subset(dfVal, select = Attrition)



```
```{r BasicFactoring}

#cleaning and munging data, persistence, statistics
#Function to convert monthly income to annual income 
MonthToYear <- function(x){
  x = x * 12
}
dfTrain$AnnualIncome <- as.integer(lapply(dfTrain$MonthlyIncome, MonthToYear))
dfVal$AnnualIncome <- as.integer(lapply(dfVal$MonthlyIncome, MonthToYear))

#Factor character variables
dfTrain <- dfTrain %>%
  mutate_if(is.character, as.factor) %>%
  select(Attrition, everything())
dfVal <- dfVal %>%
  mutate_if(is.character, as.factor) %>%
  select(Attrition, everything())

#Factor Annual Income into 4 levels
dfTrain$AnnualIncomeGroup <- cut(dfTrain$AnnualIncome, breaks = 5)

```

```{r numericfactoring}

#Factor Age of Employee into Age Groups, 5 levels
dfTrain$HourlyRate <- cut(dfTrain$HourlyRate, breaks = 5)
dfTrain$DailyRate <- cut(dfTrain$DailyRate, breaks = 5)
dfTrain$MonthlyIncome <- cut(dfTrain$MonthlyIncome, breaks = 5)
dfTrain$MonthlyRate <- cut(dfTrain$MonthlyRate, breaks = 5)
dfTrain$Age <- cut(dfTrain$Age, breaks = 5)

```

```{r factoreverything}


dfTrain$Education <- factor(dfTrain$Education)
dfVal$Education <- factor(dfVal$Education)

dfTrain$StockOptionLevel <- factor(dfTrain$StockOptionLevel)
dfVal$StockOptionLevel <- factor(dfVal$StockOptionLevel)

dfTrain$PerformanceRating <- factor(dfTrain$PerformanceRating)
dfVal$PerformanceRating <- factor(dfVal$PerformanceRating)

dfTrain$JobSatisfaction <- factor(dfTrain$JobSatisfaction)
dfVal$JobSatisfaction <- factor(dfVal$JobSatisfaction)

dfTrain$JobLevel <- factor(dfTrain$JobLevel)
dfVal$JobLevel <- factor(dfVal$JobLevel)

dfTrain$DistanceFromHome <- cut(dfTrain$DistanceFromHome, breaks = 5)
dfVal$DistanceFromHome <- cut(dfVal$DistanceFromHome, breaks = 5)

dfTrain$JobInvolvement <- factor(dfTrain$JobInvolvement)
dfVal$JobInvolvement <- factor(dfVal$JobInvolvement)

dfTrain$NumCompaniesWorked <- cut(dfTrain$NumCompaniesWorked, breaks = 5)
dfVal$NumCompaniesWorked <- cut(dfVal$NumCompaniesWorked, breaks = 5)

dfTrain$EnvironmentSatisfaction <- factor(dfTrain$EnvironmentSatisfaction)
dfVal$EnvironmentSatisfaction <- factor(dfVal$EnvironmentSatisfaction)

dfTrain$RelationshipSatisfaction <- factor(dfTrain$RelationshipSatisfaction)
dfVal$RelationshipSatisfaction <- factor(dfVal$RelationshipSatisfaction)

dfTrain$EnvironmentSatisfaction <- factor(dfTrain$EnvironmentSatisfaction)
dfVal$EnvironmentSatisfaction <- factor(dfVal$EnvironmentSatisfaction)

dfTrain$WorkLifeBalance <- factor(dfTrain$WorkLifeBalance)
dfVal$WorkLifeBalance <- factor(dfVal$WorkLifeBalance)

dfTrain$TotalWorkingYears <- cut(dfTrain$TotalWorkingYears, breaks = 5)
dfVal$TotalWorkingYears <- cut(dfVal$TotalWorkingYears, breaks = 5)

dfTrain$YearsAtCompany <- cut(dfTrain$YearsAtCompany, breaks = 5)
dfVal$YearsAtCompany <- cut(dfVal$YearsAtCompany, breaks = 5)

dfTrain$YearsInCurrentRole <- cut(dfTrain$YearsInCurrentRole, breaks = 5)
dfVal$YearsInCurrentRole <- cut(dfVal$YearsInCurrentRole, breaks = 5)

dfTrain$YearsSinceLastPromotion <- cut(dfTrain$YearsSinceLastPromotion, breaks = 5)
dfVal$YearsSinceLastPromotion <- cut(dfVal$YearsSinceLastPromotion, breaks = 5)

dfTrain$PercentSalaryHike <- cut(dfTrain$PercentSalaryHike, breaks = 5)
dfVal$PercentSalaryHike <- cut(dfVal$PercentSalaryHike, breaks = 5)

dfTrain$YearsWithCurrManager <- cut(dfTrain$YearsWithCurrManager, breaks = 5)
dfVal$YearsWithCurrManager <- cut(dfVal$YearsWithCurrManager, breaks = 5)

dfTrain$TrainingTimesLastYear <- cut(dfTrain$TrainingTimesLastYear, breaks = 5)
dfVal$TrainingTimesLastYear <- cut(dfVal$TrainingTimesLastYear, breaks = 5)

```



```{r ExploratoryDataAnalysis}
attach(dfTrain)
#Attrition Rate in dfTrain Sample Set

dfTrainColumns <- colnames(dfTrain)
print(dfTrainColumns, quote = FALSE)
dfValColumns <- colnames(dfVal)


#Finding Attrition Rates for Training Set and Character Variables
###dfTrain Sample Attrition Rate
dfTrain %>% group_by(Attrition) %>%
  count(Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by Age
dfTrain %>% group_by(Age) %>%
  count(Age, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by AnnualIncome
dfTrain %>% group_by(AnnualIncome) %>%
  count(AnnualIncome, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrtion Rate for BusinessTravel
dfTrain %>% group_by(BusinessTravel) %>%
  count(BusinessTravel, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by Department
dfTrain %>% group_by(Department) %>%
  count(Department, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)


###Attrition Rate by DistanceFromHome
dfTrain %>% group_by(DistanceFromHome) %>%
  count(DistanceFromHome, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by Education
dfTrain %>% group_by(Education) %>%
  count(Education, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by EducationField
dfTrain %>% group_by(EducationField) %>%
  count(EducationField, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by EnvironmentSatisfaction
dfTrain %>% group_by(EnvironmentSatisfaction) %>%
  count(EnvironmentSatisfaction, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by Gender
dfTrain %>% group_by(Gender) %>%
  count(Gender, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by JobInvolvement
dfTrain %>% group_by(JobInvolvement) %>%
  count(JobInvolvement, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by JobLevel
dfTrain %>% group_by(JobLevel) %>%
  count(JobLevel, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by JobRole
dfTrain %>% group_by(JobRole) %>%
  count(JobRole, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by JobSatisfaction
dfTrain %>% group_by(JobSatisfaction) %>%
  count(JobSatisfaction, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by MaritalStatus
dfTrain %>% group_by(MaritalStatus) %>%
  count(MaritalStatus, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by NumCompaniesWorked
dfTrain %>% group_by(NumCompaniesWorked) %>%
  count(NumCompaniesWorked, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by OverTime
dfTrain %>% group_by(OverTime) %>%
  count(OverTime, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by PercentSalaryHike
dfTrain %>% group_by(PercentSalaryHike) %>%
  count(PercentSalaryHike, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by PerformanceRating
dfTrain %>% group_by(PerformanceRating) %>%
  count(PerformanceRating, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by RelationshipSatisfaction
dfTrain %>% group_by(RelationshipSatisfaction) %>%
  count(RelationshipSatisfaction, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by StockOptionLevel
dfTrain %>% group_by(StockOptionLevel) %>%
  count(StockOptionLevel, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by TotalWorkingYears
dfTrain %>% group_by(TotalWorkingYears) %>%
  count(TotalWorkingYears, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by TrainingTimesLastYear
dfTrain %>% group_by(TrainingTimesLastYear) %>%
  count(TrainingTimesLastYear, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by WorkLifeBalance
dfTrain %>% group_by(WorkLifeBalance) %>%
  count(WorkLifeBalance, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by YearsAtCompany
dfTrain %>% group_by(YearsAtCompany) %>%
  count(YearsAtCompany, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by YearsInCurrentRole
dfTrain %>% group_by(YearsInCurrentRole) %>%
  count(YearsInCurrentRole, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by YearsSinceLastPromotion
dfTrain %>% group_by(YearsSinceLastPromotion) %>%
  count(YearsSinceLastPromotion, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

###Attrition Rate by YearsWithCurrManager
dfTrain %>% group_by(YearsWithCurrManager) %>%
  count(YearsWithCurrManager, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)




###Attrition Rates by Business Travel by Department
dfTrain %>% group_by(Department, BusinessTravel) %>%
  count(BusinessTravel, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)



#The Data shows that Research and Development that Travel Frequently and Sales People that Travel Frequently have the highest attrition rates

dfTrain %>% group_by(Department, BusinessTravel, JobRole, JobInvolvement, JobLevel, StockOptionLevel) %>%
  count(BusinessTravel, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

dfTrain %>% group_by(MaritalStatus, RelationshipSatisfaction, Gender) %>%
  count(Gender, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

dfTrain %>% group_by(YearsWithCurrManager, JobSatisfaction) %>%
  count(YearsWithCurrManager, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)

#Employees with low job satisfaction and a new manager are at higher risk of leaving
#If we know an employee is leaving, we should look to see if they had a new manager

dfTrain %>% group_by(YearsInCurrentRole, YearsSinceLastPromotion, JobSatisfaction) %>%
  count(YearsInCurrentRole, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)
#It appears that an employee at 5-7 years in the same role with with as many years since last promotion is at higher risk of leaving the company. 

dfTrain %>% group_by(JobSatisfaction, JobRole) %>%
  count(JobSatisfaction, Attrition) %>%
  spread(Attrition, n) %>%
  mutate(LevelAttritionRate = 100*(Yes/(Yes + No))) %>%
  mutate(SampleAttritionRate = (Yes/1170)*100)


```
