---
title: "CaseStudy2_mmastrangeli - "
author: "Mark Mastrangeli - DDS Analytics, Chief Data Scientist :)"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(base)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(knitr)
library(mlr)
library(caret)
library(recipes)
#colnames(read.csv("CaseStudy2-data.csv", header = TRUE))
```

#Don't be Penny Wise and Pound Foolish with your talent
##Remember they have real human challenges too

###The Plan
*Import
*Tidy
*Transform
*Visualize
*Model (Linear Regression) 
***How do I model a random variable like age or a scaled (1-5 or 1-30 range) variable against a binary (0:1, Yes/No)
*Communicate

##Summary: Human Capital; the talent that is the lifeblood of your orginization, especially your top talent, are your (orginizations) most valuable asset. Losing high quality, talented employees is a far bigger financial loss to the orginization than the sum of that employees total compensation package.  Said another way, it will cost FAR more to replace talented employees who meet or exceed the requirements of their role and are positive contributors to the overall good of the company, than it does to keep them. 

##Your best human capital and employee compensation liability are people too. Employees real human experiences and personal life DO come into work whether we like it or not.  


##Working to identify statistically, whether or not any of the factors and data points we have on this particular sample of employees who have come and gone is a worthy exercise if we can identify accurate predictive factors or leading indicators toward employee attrition. 

##It is my assumption that 

#Preliminary "hunches": 
*Stock Options and Pay are key factors in high grade or high job level, high involvement employee retention
**StockOptions >= 2 Is attrition rate lower than attrition rate when stock options <2
*For Tenure > 10, Turnover of tenured, joblevel > 3, high job level employees, timesincepromotion > 5 increase in time since raise or promotion, 
*Turnover of employees with long tenure but less than a year with a current manager could be interpersonal issues with new manager. 
**if attrition = true did employeeID have <2 YearsWithCurrentManager?
*Years at company vs years since promotion would seem to be a strongly correlated derived factor. 
*Lower grade or job level employees at lower income levels will have higher rates of turnover.  
**Is Attrition Rate higher when joblevel <= 2, and/or when incomelevel <=2

(All dfTrain$ID where HourlyRate && JobLevel && 

# [1] "Age"                      "Attrition"                "BusinessTravel"           "DailyRate"                "Department"              
# [6] "DistanceFromHome"         "Education"                "EducationField"           "EmployeeNumber"           "EnvironmentSatisfaction" 
#[11] "Gender"                   "HourlyRate"               "JobInvolvement"           "JobLevel"                 "JobRole"                 
#[16] "JobSatisfaction"          "MaritalStatus"            "MonthlyIncome"            "MonthlyRate"              "NumCompaniesWorked"      
#[21] "OverTime"                 "PercentSalaryHike"        "PerformanceRating"        "RelationshipSatisfaction" "StandardHours"           
#[26] "StockOptionLevel"         "TotalWorkingYears"        "TrainingTimesLastYear"    "WorkLifeBalance"          "YearsAtCompany"          
#[31] "YearsInCurrentRole"       "YearsSinceLastPromotion"  "YearsWithCurrManager"     "AnnualIncome"             "AnnIncomeLevel"          
#[36] "AgeGroup"                 "JobChanges" 

colnames(dfTrain)

#Key Factors that lead to attrition or employee turnover

##The goal of this project is to identify key factors that lead to employee attrition in order to provide the highest quality talent with the lowest probability of churn to our clients. 
##No Company wants to invest in the onboarding and deveopment of a new employee and have that employee leave the company after a return can be made on the investment. Nor does any company want to have a work environment that creates risk of loss of talent that is the life blood of the orginization. 

#The Key question we are seeking to answer is whether or not Data Science and Analytics can help us identify talent with a lower risk of attrition for our clients. 

Collecting, cleaning and munging data
Using persistence, statistics and software engineering skills
Understanding biases in data
Debugging output from code
Combining visualization and data sense to explor data
Finding patterns, building models, writing algorithms
Devising ways to use data as a product or to improve a product
Designing experiments
Communicating with team members

*1. Basic Analysis of the data - 
*2. Theories and Hunches
*3. Observations and Analysis of given variables (later, potentially observations and analaysis of derived variables)
*4.

*Import
*Tidy
*Transform
*Visualize
*Model (Linear Regression) 
***How do I model a random variable like age or a scaled (1-5 or 1-30 range) variable against a binary (0:1, Yes/No)
*Communicate

```{r, Import}
#Read in CSV and change columns of interest to factors

dfVal <- read_csv("CaseStudy2Validation.csv")
dfTrain <- read_csv("CaseStudy2-data.csv")

```

```{r}

#Taking a look at the data to see what we are working with!
#str(dfTrain)
# We have 37 variables and 1170 observations. The Variable columns seem to have the correct variable object types. 

#dfTrain <- subset(dfTrain, select = -c(Rand, EmployeeCount, Over18)) #Dropping Employee ID and Rand arbitrary variables as well as variables with only 1 level: EmployeeCount, Over18
summary(dfTrain)



#Factor

#Add column of annual income data by multiplying monthly income by 12
#Function to convert monthly income to annual income 
MonthToYear <- function(x){
  x = x * 12
}
dfTrain$AnnualIncome <- as.integer(lapply(dfTrain$MonthlyIncome, MonthToYear))

#Factor character variables
dfTrain <- dfTrain %>%
  mutate_if(is.character, as.factor) %>%
  select(Attrition, everything())

#Factor Annual Income into 6 levels
dfTrain$AnnualIncome <- cut(dfTrain$AnnualIncome, breaks=c(0,50000,100000,199999,Inf),labels=c("$0-$50,000", "$50,000-$100,000", "$100,000-$200,000", "Over $200,000"))

#Factor Age of Employee into Age Groups, 8 levels
dfTrain$AgeGroup <- cut(dfTrain$Age, breaks = c(18,20,22,25,30,35,40,45,50,55,60,Inf), labels = c("18-20", "20-22", "22-25", "25-30", "30-35", "35-40", "40-45", "45-50", "50-55", "55-60", "60+"))

dfTrainAttritionDummies <- model.matrix(~dfTrain$Attrition -1, data = dfTrain)
dfTrainBusinessTravelDummies <- model.matrix(~dfTrain$BusinessTravel -1, data = dfTrain)
dfTrainEducationFieldDummies <- model.matrix(~dfTrain$Attrition -1, data = dfTrain)
dfTrainGenderDummies <- model.matrix(~dfTrain$Gender -1, data = dfTrain)
dfTrainDepartmentDummies <- model.matrix(~dfTrain$Department -1, data = dfTrain)
dfTrainEducationFieldDummies <- model.matrix(~dfTrain$EducationField -1, data = dfTrain)
dfTrainJobRoleDummies <- model.matrix(~dfTrain$JobRole -1, data = dfTrain)
dfTrainJobSatisfactionDummies <- model.matrix(~dfTrain$JobSatisfaction -1, data = dfTrain)
dfTrainMaritalStatusDummies <- model.matrix(~dfTrain$MaritalStatus -1, data = dfTrain)
dfTrainOverTimeDummies <- model.matrix(~dfTrain$OverTime -1, data = dfTrain)
dfTrainAnnualIncomeDummies <- model.matrix(~dfTrain$AnnualIncome -1, data = dfTrain)
dfTrainAgeGroupDummies <- model.matrix(~dfTrain$AgeGroup -1, data = dfTrain)

lm(formula = dfTrainAttritionDummies ~ dfTrainBusinessTravelDummies + dfTrainEducationFieldDummies + dfTrainGenderDummies + dfTrainDepartmentDummies + dfTrainEducationFieldDummies + dfTrainJobRoleDummies + dfTrainJobSatisfactionDummies + dfTrainMaritalStatusDummies + dfTrainOverTimeDummies + dfTrainAnnualIncomeDummies)

#Based on this model, it appears that the BusinessTravel_Frequently and DepartmentResearch&Development have the most positive relationship to the attritionYes outcome.


```

```{r}
dfTrainAttrition = filter(dfTrain, Attrition == "Yes")
dfTrainAttritionRate = (count(dfTrainAttrition))/1170
colnames(dfTrainAttritionRate) <- c("dfTrainAttritionRate")
print(dfTrainAttritionRate)


#Attrition Rate Per Variable Factor Level
#Attrition Rate by Job Satisfaction


#How do I create a function to do this for every variable and level
#dfTrain.var <- c(colnames(dfTrain$4)
#variable.names(dfTrain)
#dfTrain.var = filter(dfTrain, dfTrain.var >= 0 )



#Filter ob Satisfaction Factor Level 1
dfTrain.JobSatisfactionL1 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 1)
#Calculate Job Satisfaction Level 1 Attrition Rate
dfTrain.JobSatAttrRateL1 = count(dfTrain.JobSatisfactionL1)/1170
#Factor Level 2
dfTrain.JobSatisfactionL2 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 2)
dfTrain.JobSatAttrRateL2 = count(dfTrain.JobSatisfactionL2)/1170
#Factor Level 3
dfTrain.JobSatisfactionL3 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 3)
dfTrain.JobSatAttrRateL3 = count(dfTrain.JobSatisfactionL3)/1170
#Factor Level 4
dfTrain.JobSatisfactionL4 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 4)
dfTrain.JobSatAttrRateL4 = count(dfTrain.JobSatisfactionL4)/1170

#Attrition Rate of Frequent Business Travelers
dfTrain.BusinessTravelL3 = filter(dfTrain, Attrition == "Yes" & BusinessTravel == "Travel_Frequently")
dfTrain.BusinessTravelAttrRateL3 = count(dfTrain.BusinessTravelL3)/1170
#Attrition Rate of Employees in Research and Development
dfTrain.DepartmentL1 = filter(dfTrain, Attrition == "Yes" & Department == "Research & Development")
dfTrain.DepartmentL1AttrRate = count(dfTrain.DepartmentL1)/1170

contrasts(dfTrain$Department)




#Job Satisfaction Level 1 Attrition Rate
dfTrain.JobSatAttrRateL1 = count(dfTrain.JobSatisfactionL1)/1170
#Factor Level 2
dfTrain.JobSatisfactionL2 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 2)
dfTrain.JobSatAttrRateL2 = count(dfTrain.JobSatisfactionL2)/1170
#Factor Level 3
dfTrain.JobSatisfactionL3 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 3)
dfTrain.JobSatAttrRateL3 = count(dfTrain.JobSatisfactionL3)/1170
#Factor Level 4
dfTrain.JobSatisfactionL4 = filter(dfTrain, Attrition == "Yes" & JobSatisfaction == 4)
dfTrain.JobSatAttrRateL4 = count(dfTrain.JobSatisfactionL4)/1170


```




```{r}
ggplot(dfTrain,aes(x=factor(Attrition),fill=factor(JobSatisfaction)))+
   geom_bar(position="dodge")

ggplot(dfTrain,aes(x=factor(JobSatisfaction), y=factor(Attrition),fill=factor(JobSatisfaction)))+
 geom_jitter()
ggplot(dfTrain,aes(x=factor(JobSatisfaction), y=factor(Attrition),fill=factor(JobSatisfaction)))+
 geom_point()

ggplot(dfTrain,aes(x=factor(JobInvolvement),fill=factor(Attrition)))+
  geom_bar(position="dodge")+
  facet_grid(". ~ Attrition")
ggplot(dfTrain,aes(x=factor(YearsSinceLastPromotion),fill=factor(Attrition)))+
  geom_bar(position="dodge")+
  facet_grid(". ~ Attrition")
ggplot(dfTrain,aes(x=factor(YearsSinceLastPromotion),fill=factor(Attrition)))+
  geom_bar(position="dodge")+
  facet_grid(". ~ Attrition")

ggplot(dfTrain, aes(x=Attrition,  fill=Attrition))+
labs(title="Employees", x = "", y="Count", fill="Attrition") + geom_bar(stat = "count")

ggplot(dfTrain, aes(x=Attrition,  fill=Attrition))+
labs(title="Employees", x = "", y="Count", fill="Attrition") + geom_bar(stat = "count")

ggplot(data = dfTrain) +
  geom_bar(mapping = aes(x=AnnualIncome))

ggplot(dfTrain, aes(x=AnnualIncome, fill = Attrition)) + labs(title="Attrition Frequency by Annual Income", x = "Annual Income", y="Count", fill="Attrition") + scale_fill_brewer("Dark")

ggplot(dfTrain, aes(x=NumCompaniesWorked,fill = Attrition)) + 
  labs(title="Employee Distribution by Number of Companies Worked", x = "Number of Companies", y="Employee Count", fill="Status") +
  geom_histogram( binwidth=1, position = position_stack(reverse = TRUE),color="white") +
  scale_fill_manual(values=c("#2F63A0","#7D9099"))

#relabel  x axis, show actual count on bars. Show rate on bar per level

plot(HourlyRate ~ Attrition, data = dfTrain, xlab = "Attrition Rate", ylab = "Annual Income")
plot(StockOptionLevel ~ Attrition, data = dfTrain, xlab = "Attrition Rate", ylab = "Stock Option Level")
plot(TotalWorkingYears ~ Attrition, data = dfTrain, xlab = "Attrition Rate", ylab = "Years at Company")
plot(YearsInCurrentRole ~ Attrition, data = dfTrain, xlab = "Attrition Rate", ylab = "Years in Current Role")
plot(YearsWithCurrManager ~ Attrition, data = dfTrain, xlab = "Attrition Rate", ylab = "Years with Current Manager")
plot(YearsAtCompany ~ Attrition, data = dfTrain, xlab = "Attrition Rate", ylab = "Years at Company")
```



```{r, InitialExploration}








ggplot(dfTrain, aes(x=Attrition,  fill=Attrition))+
labs(title="CaseStudy2_data Attrition Frequency", x = "", y="Count", fill="Attrition") + geom_bar(stat = "count") + scale_fill_brewer(palette = "Dark2")

#What is highest or most common variable on sample which has left the company?

ggplot(dfTrain, aes(x=AnnualIncome,fill = Attrition))+
  labs(title="Employee Distribution by Hourly Rate", x = "Annual Income", fill="AnnualIncome") +
geom_histogram(binwidth=7, color="white",position = position_stack(vjust = 1, reverse = FALSE)) +
  scale_fill_brewer(palette = "Dark2")

ggplot(dfTrain, aes(x=HourlyRate,fill = Attrition))+
  labs(title="Employee Distribution by Hourly Rate", x = "Hourly Rate", y="Employee Count", fill="Status") +
geom_histogram(binwidth=7, color="white",position = position_stack(vjust = 1, reverse = FALSE)) +
  scale_fill_brewer(palette = "Dark2")

dfTrain %>% ggplot(aes(y = Attrition, fill = Attrition)) + facet_wrap(dfTrain$BusinessTravel, dfTrain$JobSatisfaction)
  labs(title="Employee Distribution by Hourly Rate", x = "Hourly Rate", y="Employee Count", fill="Status") +
geom_histogram(binwidth=7, color="white",position = position_stack(vjust = 1, reverse = FALSE)) +
  scale_fill_brewer(palette = "Dark2")
beerCOTX %>% ggplot(aes(x = IBU, y = ABV)) + facet_wrap(~State) + geom_point() + labs(title = "ABV v. IBU for Colorado")

dfTrain %>% 
  filter(Attrition == "Yes") 
dfTrain %>% 
  select(BusinessTravel) #How do I select variables where the value is over 2 or 3?

#plot(filter(beerCOTX,State == "TX")$IBU,filter(beerCOTX,State == "TX")$ABV, pch = 20, col = "blue", xlab = "International Bitterness Units", ylab = "Alcohol By Volume", main = "ABV v. IBU of Texas Beers") lines(filter(beerCOTX,State == "TX")$IBU,predsTX)

#ggplot(VAMC_State_Count, aes(reorder(VAMC_State_Count$State, VAMC_State_Count$`Number of VAMC`), VAMC_State_Count$`Number of VAMC`)) + geom_bar(aes(fill=VAMC_State_Count$State), stat="identity") + ggtitle("Distribution of VA Medical Centers in continental US") + theme(axis.text.x=element_text(angle = 90, vjust = 1,hjust = ), legend.position="none") + scale_color_manual(values = LST) + labs(title="Number of Veterans Affairs Medical Centers by State", fill = "State", x= "State", y= "VAMC Count")

```




```{r models}

dfTrain %>%
  group_by(Attrition, JobInvolvement, JobSatisfaction) %>%
  summarize(
    #Average Job Involvement
    count(JobInvolvement),
    #Average Job Satisfaction
    count(JobSatisfaction)
  )


```

```{r}

#Is there a different percentage of Male or Female attrition? 
#Sex of Employee: Factor with Two levels, "Female"(2), "Male"(1)
GenderCount <- count(dfTrain, dfTrain$Gender)
#NumberOfWomen <- as.integer(GenderCount[2,2])
#NumberOfMen <- as.integer(GenderCount[1,2])

MaleAttritionNormPercentage
FemaleAttritionNormPercentage




#ow does financial compensation influence attrition?
HourlyRateRange <- range()
AvergageHourlyRate
MaxHourlyRate
MedianHourlyRate
MeanHourlyRate
MeanHourlyRateMen
MeanHourlyRateWomen
MaxHourlyRateMen
MaxHourlyRateWomen

#How does workload, worklife balance, business travel, distance from home, job involvement, overtime, job level effect the attrition rate?

#Non-TravelAttritionRate
#Travel_RarelyAttritionRate
#Travel_FrequentlyAttritionRate
#MaxDistanceFromHome
#AvgDistanceFromHome





```

```{r}
#Biomass Recipe Example: 
head(dfTrain)
head(dfVal)


rec <- recipe(Attrition ~ Age + AnnualIncome + BusinessTravel + DailyRate + Department + DistanceFromHome + Education + EducationField + EmployeeNumber + EnvironmentSatisfaction + Gender + HourlyRate + JobInvolvement + JobLevel + JobSatisfaction +  RelationshipSatisfaction + MaritalStatus + MonthlyIncome + MonthlyRate + NumCompaniesWorked + OverTime + PercentSalaryHike + PerformanceRating + RelationshipSatisfaction + StandardHours + StockOptionLevel + TotalWorkingYears + TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager,
              data = dfTrain)

# Age, Attrition, BusinessTravel, DailyRate, Department, DistanceFromHome, Education, EducationField, EmployeeNumber, EnvironmentSatisfaction, Gender, HourlyRate,      "JobInvolvement"           "JobLevel"                 "JobRole"   
#[16] "JobSatisfaction"          "MaritalStatus"            "MonthlyIncome"            "MonthlyRate"              "NumCompaniesWorked"      
#[21] "OverTime"                 "PercentSalaryHike"        "PerformanceRating"        "RelationshipSatisfaction" "StandardHours"           
#[26] "StockOptionLevel"         "TotalWorkingYears"        "TrainingTimesLastYear"    "WorkLifeBalance"          "YearsAtCompany"          
#[31] "YearsInCurrentRole"       "YearsSinceLastPromotion"  "YearsWithCurrManager AnnualIncome AnnIncomeLevel AgeGroup JobChanges

center_trans <- rec %>%
  step_center(carbon, contains("gen"), -hydrogen)

center_obj <- prep(center_trans, training = biomass_tr)

transformed_te <- bake(center_obj, biomass_te)

biomass_te[1:10, names(transformed_te)]
transformed_te

tidy(center_trans, number = 1)
tidy(center_obj, number = 1)

```


```{r knn}



nTrain.points <- 1170
nVal.points <- 300

num.test.set.labels <- nTrain.points * (nTrain.points - nVal.points)

training <- sample(1:nTrain.points, nVal.points, replace = FALSE)

train <- subset(dfTrain[training, ], select = c(Age, HourlyRate))

testing <- setdiff(1:nTrain.points, training)

test <- subset(dfTrain[testing,], select = c(Age, HourlyRate))

cl <- dfTrain$Attrition[training]

true.labels <- dfTrain$Attrition[testing]


knn3Train(train, test, cl, k=9)

for(k in 1:20) {
  print(k)
  predicted.labels <- knn3Train(train, test, cl, k)
  num.incorrect.labels <- sum(predicted.labels != true.labels)
  misclassification.rate <- num.incorrect.labels / num.test.set.labels
  print(misclassification.rate)
}

test <- c(50,20)
knn3Train(train,test,cl,k=5)
  
```


```{r}
#Linear Regression
sp <- ggplot()


```

```{r}
#write.csv(Case2PredictionsSadler.csv)
dfTrainPredictions
dfTestPredictions <- 
```



<!--
```{r KMeansAndKNN}
summarizeColumns(dfTrain)
ldfTrain = length(dfTrain$ID)
ldfVal = length(dfVal$ID)

ldfTrain = length(dfTrain$ID)
ldfVal = length(dfVal$ID)

perc_test = ldfVal / (ldfVal + ldfTrain)
perc_train = ldfTrain / (ldfVal + ldfTrain)
perc_test
perc_train

head(dfTrain)
head(dfVal)

summarizeColumns(dfTrain)
summarizeColumns(dfVal)

nTrain.points <- 1170
nVal.points <- 300

perc_test <- ldfTrain * (ldfTrain - ldfVal)

training <- sample(1:ldfTrain, ldfVal, replace = FALSE)

train <- subset(dfTrain[training, ], select = c(Department, BusinessTravel))

testing <- setdiff(1:ldfTrain, training)

test <- subset(dfTrain[testing,], select = c(Department, BusinessTravel))

cl <- dfTrain$Attrition[training]

true.labels <- dfVal$Attrition[testing]


knnregTrain(train, test, cl, k=9)


#knn using the class package
#Note... you cannot put the repsonse in teh dfTrain and dfVal arguments.  
results <- knnregTrain(train, test, cl, k = 3)
dfVal$AttritionPred = results
confusionMatrix(table(dfVal$Attrition,dfVal$AttritionPred))


### KNN Regression

plot(dfTrain$Department,dfTrain$BusinessTravel, main = "Simple KNN Example")
text(dfTrain$x1, dfTrain$x2, labels = dfTrain$y, pos = 4, col = "blue")
dfVal = data.frame(x1 = dfVal$JobRole, x2 = dfVal$Department, y = as.numeric(dfVal$Attrition))




# K Means Clustering 

clusters = kmeans(dfTrain, centers = 3, nstart = 10)
dfTrain$clusters = clusters$cluster
dfTrain %>% ggplot(aes(x = x1, y = x2, label = clusters)) + geom_text()

# Classification using Kmeans clustering and KNN

dfVal = data.frame(x1 = c(6,7,8,8), x2 = c(9,10,10,11), y = c(8,9,9,7), clusters = c(3,1,1,2))
results = class::knn(dfTrain[,c(1:3)],dfVal[,c(1:3)],dfTrain[,4], k = 3)
dfVal$clustersPred = results
table(dfVal$clusters,dfVal$clustersPred) # function confusionMatrix only for binary classfication.

summarizeColumns(dfTrain)
ldfTrain = length(dfTrain$ID)
ldfVal = length(dfVal$ID)

ldfTrain = length(dfTrain$ID)
ldfVal = length(dfVal$ID)

perc_test = ldfVal / (ldfVal + ldfTrain)
perc_train = ldfTrain / (ldfVal + ldfTrain)
perc_test
perc_train

head(dfTrain)
head(dfVal)

summarizeColumns(dfTrain)
summarizeColumns(dfVal)

nTrain.points <- 1170
nVal.points <- 300

perc_test <- ldfTrain * (ldfTrain - ldfVal)

training <- sample(1:ldfTrain, ldfVal, replace = FALSE)

train <- subset(dfTrain[training, ], select = c(Department, BusinessTravel))

testing <- setdiff(1:ldfTrain, training)

test <- subset(dfTrain[testing,], select = c(Department, BusinessTravel))

cl <- dfTrain$Attrition[training]

true.labels <- dfVal$Attrition[testing]


knnregTrain(train, test, cl, k=9)


#knn using the class package
#Note... you cannot put the repsonse in teh dfTrain and dfVal arguments.  
results <- knnregTrain(train, test, cl, k = 3)
dfVal$AttritionPred = results
confusionMatrix(table(dfVal$Attrition,dfVal$AttritionPred))


### KNN Regression

plot(dfTrain$Department,dfTrain$BusinessTravel, main = "Simple KNN Example")
text(dfTrain$x1, dfTrain$x2, labels = dfTrain$y, pos = 4, col = "blue")
dfVal = data.frame(x1 = dfVal$JobRole, x2 = dfVal$Department, y = as.numeric(dfVal$Attrition))




# K Means Clustering 

clusters = kmeans(dfTrain, centers = 3, nstart = 10)
dfTrain$clusters = clusters$cluster
dfTrain %>% ggplot(aes(x = x1, y = x2, label = clusters)) + geom_text()

# Classification using Kmeans clustering and KNN

dfVal = data.frame(x1 = c(6,7,8,8), x2 = c(9,10,10,11), y = c(8,9,9,7), clusters = c(3,1,1,2))
results = class::knn(dfTrain[,c(1:3)],dfVal[,c(1:3)],dfTrain[,4], k = 3)
dfVal$clustersPred = results
table(dfVal$clusters,dfVal$clustersPred) # function confusionMatrix only for binary classfication.


```
-->




